---
title: "Output"
author: "Tom Alig"
date: "December 27, 2019"
output: html_document

---

```{r global, include=FALSE}
library(here)
#library(tidyverse)
#library(lubridate)
#library(scales)
#library(hms)
#library(lm)
```


```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```


```{r setup, include = FALSE}
#knitr::opts_knit$set(root.dir = '../')
```


```{r load_scripts, include = FALSE}

#source("RunningProjectR/scripts/my_track_scripts.R")
#source("RunningProjectR/run-all.R/analyses-track.R")

#source("Running/RunningProjectR/scripts/my_track_scripts.R")

################
# This one works!!
#source(here::here("RunningProjectR", "scripts", "my_track_scripts.R"))
################


#source(here::here("RunningProjectR", "run-all.R", "analyses-track2.R"))
source(here::here("RunningProjectR", "analyses-track2.R"))

#source("scripts/my_track_scripts.R")
#source("run-all.R/analyses-track.R")

#source(here("RunningProjectR", "scripts", "my_track_scripts.R"))
#source(here("RunningProjectR", "run-all.R", "analyses-track.R"))

#source(here("scripts", "my_track_scripts.R"))
#source(here("run-all.R", "analyses-track.R"))

#source("C:/Users/TomAlig06.15/Documents/R/Running/RunningProjectR/scripts/my_track_scripts.R")

#source("C:/Users/TomAlig06.15/Documents/R/Running/RunningProjectR/run-all.R/analyses-track.R")

```


For `r event`, the data look at those who ran a `r slowest.time` or better. This represents approximately the top 40% of all performances in the `r event`.  

Data compiled from athletic.net. Includes all athletes in the outdoor track seasons in the years 2011 - 2019.  

This project analyzes high school track performances over time, and relates your athlete's performances to historical performances. This is purely math-driven. The data below cannot and does not measure who was injured, who was healthy, who trained well, who trained poorly, who had off-the-track issues, who did not.  


## Athlete: `r sn`
### Grade in 2019-2020: `r grade.2019.20`
### School: `r hsteam`
#### Event: `r event`

#### Athlete's best time in `r event` in high school career, by grade:


### 1. Overview of `r event` over time:

#### A. Heat Map of performances. Will add up to 100% going across, for each grade level.  

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(HeatMap800f)
```


```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(heatmap2)
```


#### B. Mean and Median performances over time, among all thoses who had a `r slowest.time` or better.  

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(median.mean.graph)
```

#### C. Boxplot distribution of performances over time, among all those who had a `r slowest.time` or better.
##### The shaded areas for each grade are the middle 50%. A short shaded area means the middle 50% are pretty close to each other; converselty, a tall shaded area means  the middle 50% are spread apart by a lot.  
##### The lines in the middle of the shaded area show the median
##### The lines below the shaded area show the distribution of performances in the top 25%

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(boxplot.by.grade)
```

Distribution  of all performances in `r event` over time:  

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(Histogram.All)
```

Grade `r last.year.grade` distribution of all performances in `r event`:  

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(Hist.Last.Year)
```


### 2. Improvments

By how much does an athlete improve in the `r event`? This shows historical distribution of improvements from grade to grade:


```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(improve.dist)
```

### 3. College Comparisons

#### A. How well did college athletes in `r event` perform when they were in high school? This chart looks at Division 1, Division 2, and Division 3 college athletes and their performances in the `r event` when they were in high school.

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(med.hs.times.for.coll.athletes)
```


#### B. All performances of college athletes in `r event` when they were in high school:

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(scatterplot.hs.perf)
```

#### C. Adding in the Athlete's High School for comparison sake:

```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(facet.boxplots)
```


## 4. PREDICTIONS

Based purely on the algorithm. Does not factor in any injury, illness, training, or other factors that may be unique to your athlete.

I hesitate to put predictions into this report. There is enough pressure on high school kids today. They don't need some algorithm boxing them into a certain performance level. There are a million reasons why your athlete may far surpass this prediction. Likewise, there are a million reasons why your athlete may fall short of this prediction. 

