---
title: "title goes here"
#author: "Tom Alig"
#date: "December 31, 2019"
output: html_document
---

```{r global, include=FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(scales)
library(hms)
library(gridExtra)
library(grid)
#library(lm)
```


```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```


```{r setup, include = FALSE}
#knitr::opts_knit$set(root.dir = '../')
```


```{r load_scripts, include = FALSE}

# This one works!!
source(here::here("RunningProjectR", "scripts", "my_track_scripts.R"))
source(here::here("RunningProjectR", "run-all.R", "analyses-track.R"))
#source(here::here("RunningProjectR", "run-all.R", "glm_prod.R"))
```


```{r, echo = FALSE, include = FALSE}

ppredict <- read_csv(here::here("RunningProjectR", "data-output", "2019-12-31_hs_girls_800m_predictions_linear-regression_2011-19.csv"))

#predict_file <- here("RunningProjectR", "data-output", #"2019-12-31_hs_girls_800m_predictions_linear-regression_2011-19.csv")

likelihood <- read_csv(here::here("RunningProjectR", "data-output", "2019-12-31_hs_girls_800m_probabilities_glm_2011-19.csv"))

```

## **INTRODUCTION**  

This project came about as a result of my love of numbers, and my two daughters’ love of track. I had a handful of questions I kept asking myself. Are my daughters capable of being a Division 1, Division 2, or Division 3 track athlete? How did college athletes at the D1, D2, and D3 levels perform when they were in high school? How do my daughters’ performances compare to the high school performances of those athletes?  

In addition, I would wonder about their progression in their chosen events from year-to-year. What was a “normal” expectation for a change in performance from one grade to the next?  

Moreover, I thought, “What is a realistic expectation for their performances in their events this coming season? How likely is it that my daughter will reach certain performance milestones?”  

This project attempts to answer those questions, and more.  

Unlike cross-country, one track isn’t hilly, the next one isn’t muddy, the third one isn’t run over branches and other obstacles in the ground. The numbers/comparison geek in me likes the fact that, for the most part, tracks are relatively equal to compare. Sure the wind might kick up one day, or one track surface might be in great condition and another in poor condition, but by and large the measurement part of it is done under more consistent conditions than cross country.  

I gathered high school data from athletic.net for the years `r years`. College data were gathered from a variety of free online sources, including league and individual school websites, for the years 2018 and 2019. I then set out to answer the questions above, and more.  


<br>  


*****  


## **REPORT**  

### Athlete: `r sn`
#### Grade in 2019-2020: `r grade.2019.20`
#### School: `r hsteam`
#### Event: `r event2`

#### Athlete's best time in `r event` in high school career, by grade:  
#### Grade `r last.year.grade`: `r athlete.data[3,2]`
#### Grade `r two.years.ago.grade`: `r athlete.data[2,2]`


<br>  


*****  



## **1. BIG PICTURE**  

The high school data in this represents that of all athletes who ran a `r slowest.time` or better in the years `r years` in the outdoor track season. Most athletes did NOT run this well. If your athlete ran a `r slowest.time` then they are among the top 40% of all performances over time in the `r event2`!   

It’s hard to be a high school varsity athlete. It’s even harder to achieve the threshold listed above. Keep this in mind as you compare your athlete’s performances. A `r slowest.time` or better is a very good performance!  

Keep in mind that the performances analyzed here are the best performance for each athlete, each year. Your athlete may compete in a dozen or more track meets: only their best performance in each event is analyzed here.  


<br>  


*****  



## **2. DATA OVERVIEW**    

How many athletes run a `r slowest.time` or better each year in the `r event2`? Figure 1 shows the number of runners by year. A factor in the increasing number of athletes in this event is merely the availability of data. Ten years ago some schools had their information readily available. Now most schools do, and as a result we see higher participation numbers when it may simply be the case that we now have more data available.  


<br>  


Figure 1  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(num.of.participants)
```


<br>  


*****  



## **3. DISTRIBUTIONS**    

Within the parameter of performances that are `r slowest.time` or better, what is considered an average time? What constitutes an excellent performance? A very good performance? A good performance?  

Figure 2 shows The Median and Mean performances for the `r event2`, among those who ran a `r slowest.time` or better. Remember that median is the middle value, and the mean is the average of all values.  


<br>  


Figure 2  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(median.mean.barplot)
```


<br>  


*****  



Distribution charts can also help us see the big picture. What are the absolute best performances? How common are they? Figure 3 shows the distribution of all performances over time.  


<br>  


Figure 3  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(Histogram.All)
```


<br>  


*****  



How do the performances vary by grade? The heatmap in Figure 4 helps to show grade-level variability. The numbers will add up to 100% going **across**, to represent 100% for each grade.  
The orange-colored square depicts the performance of your athlete.  


<br>  


Figure 4  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
plot(heatmap2)
```


<br>  


*****  



How do the distributions change by grade level? Are seniors performing better than freshmen? By how much?  
The boxplot in Figure 5 shows some differences.  

* The shaded areas for each grade are the middle 50% of performances for that grade.  

* A short shaded area means the middle 50% are pretty close to each other; conversely, a tall shaded area means there is a wide range in the middle 50%.  

* The solid lines in the middle of the shaded areas show the median for each grade.  

* The lines below the shaded area show the distribution of performances in the top 25%.  

* Closer to the “best” performances you will notice some dots. This indicates that there are very few results at these performance levels. These are the outliers, the excellent performances for `r event`.  


<br>  


Figure 5  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(boxplot.by.grade)
```


<br>  


*****  



## **4. IMPROVEMENT**  

Keep in mind these athletes are already performing at a high level. The better your initial performance, the less your improvement will be. An athlete who runs a four-minute mile can only marginally improve their time, while the athlete who runs a ten-minute mile can perhaps improve their time by a significant amount.  

Figure 6 depicts the historical distribution of improvements from grade to grade.  


<br>  


Figure 6  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(improve.dist)
```


<br>  


*****  



##  **5. COLLEGE COMPARISON**  


How well did college athletes in `r event2` perform when they were in high school? Figure 7 looks at Division 1, Division 2, and Division 3 college athletes and their performances in the `r event` when they were in high school.  


<br>  


Figure 7  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(scatterplot.hs.perf2)
```


<br>  


*****  



Taking it a step further, how did the collegiate All-Conference and All-American athletes in this event perform when they were in high school? Figure 8 shows the high school performances of future Division 1, 2, and 3 elite athletes in `r event`.  


<br>  


Figure 8  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(scatterplot.allconf.aa)
```


<br>  


*****  



The boxplot in Figure 9 show the distribution of performances for D1, D2, and D3 athletes when they were in grades 9 - 12. It also includes a distribution of performances for your athlete's school in the years `r years`.


<br>  


Figure 9  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(facet.boxplots)
```


<br>  


*****  



Finally, what were the typical performances, by high school grade, of those who went on to be collegiate track and field athletes? Figure 10 sheds some insight.  


<br>


Figure 10  
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(1106)
plot(med.hs.line.plot)
```


<br>  


*****  



## **6. PREDICTIONS**  

Based solely on the data, what is your athlete predicted to run next season as a `r current.school.year.grade`th grader?  

I really hesitate to provide a prediction. The high school years can be fraught with enough anxiety as it is, without some random guy on the internet saying that your athlete is predicted to perform at a certain level.    

Disclaimers. There are a million reasons why your athlete may far surpass the prediction below, and there are a million reasons why your athlete may not achieve the prediction.  

The prediction(s) below do not and cannot measure:  

* If an athlete was injured last year and therefore did not perform well  

* If an athlete is injured this year and therefore cannot perform as well as last year  

* The fact that puberty comes at different grades and ages  

* The quality of an athlete's training  

* The quantity of an athlete's training  

* Factors in personal lives of athletes that may affect performance  

* Nutrition levels of an athlete  

* Many other factors  

Please take the prediction with a grain of salt.  

I used a linear regression model to predict future performances. This model performed better than other machine learning methodologies, including random forest, on training data.  


```{r, echo = FALSE, warning = FALSE, message=FALSE, results = "hide"}
#predict2 <- read_csv(predict_file)

my.pred <- ppredict %>% 
  filter(Athlete == sn) %>% 
  mutate(Gr10pred = round(as.hms(lmpred10 * 60),2)) %>%
  mutate(Gr10pred2 = as.hms(Gr10pred)) %>% 
  #mutate(Gr10pred = as.hms(lmpred10 * 60)) %>%
  mutate(Gr11pred = round(as.hms(lmpred11 * 60),2)) %>%
  mutate(Gr11pred2 = as.hms(Gr11pred)) %>% 
  mutate(Gr12pred = round(as.hms(lmpred12 * 60),2)) %>%
  mutate(Gr12pred2 = as.hms(Gr12pred)) %>% 
  select(Gr10pred2, Gr11pred2, Gr12pred2)

current.year.pred <- ifelse(current.school.year.grade == 9, "na", ifelse(current.school.year.grade == 10, as.hms(my.pred$Gr10pred2),
                            ifelse(current.school.year.grade == 11,
                                   as.hms(my.pred$Gr11pred2), 
                                   as.hms(my.pred$Gr12pred2))))

current.year.pred2 <- as.hms(current.year.pred)
current.year.pred2

wtf.pred <- ifelse(current.school.year.grade == 9, "na", ifelse(current.school.year.grade == 10, my.pred[1],
                            ifelse(current.school.year.grade == 11,
                                   my.pred[2], 
                                   my.pred[3])))

#my.pred2 <- as.hms(my.pred * 60)
#print(my.pred)
#print(my.pred$Gr10pred2)
#print(my.pred$Gr11pred2)
#print(c(my.pred$Gr10pred2, my.pred$Gr11pred2, my.pred$Gr12pred2))
```


By the numbers, and only by the numbers, the prediction for `r sn` in Grade `r grade.2019.20` `r event` is:  

### `r wtf.pred`  


<br>  


*****  



## **7. PROBABILITIES**  

I wanted to take the prediction a step further to determine the likelihood of achieving certain levels of performance. What is the probability your athlete being able to incrementally improve their performance?  

I took the data from `r years` and ran a logistic regression model for various levels of performance in the `r event2`. As shown in Figure 11, the model then calculated the likelihood of your athlete achieving these levels.  


<br>  


Figure 11   
```{r, echo = FALSE, warning = FALSE, message=FALSE, results = 'hide'}
# Gauge Plot
df_sub <- likelihood %>% 
  filter(Athlete == sn)

gg.gauge_percent <- function(pos,breaks=c(0,30,70,100)) {
  require(ggplot2)
  get.poly <- function(a,b,r1=0.5,r2=1.0) {
    th.start <- pi*(1-a/100)
    th.end   <- pi*(1-b/100)
    th       <- seq(th.start,th.end,length=100)
    x        <- c(r1*cos(th),rev(r2*cos(th)))
    y        <- c(r1*sin(th),rev(r2*sin(th)))
    return(data.frame(x,y))
  }
  ggplot()+
    geom_polygon(data=get.poly(breaks[1],breaks[2]),aes(x,y),fill="#D55E00")+
    geom_polygon(data=get.poly(breaks[2],breaks[3]),aes(x,y),fill="#F0E442")+
    geom_polygon(data=get.poly(breaks[3],breaks[4]),aes(x,y),fill="#009E73")+
    geom_polygon(data=get.poly(pos-1,pos+1,0.2),aes(x,y))+
    geom_text(data=as.data.frame(breaks), 
              size=4, 
              fontface="bold", 
              vjust=0,
              aes(x=1.1*cos(pi*(1-breaks/100)),
                  y=1.1*sin(pi*(1-breaks/100)),
                  label=paste0(breaks,"%")))+
    annotate("text",x=0,y=0,
             label=paste(pos,"%"),
             vjust=0,
             size=5,
             fontface="bold")+
    coord_fixed()+
    #ggtitle(df_sub$Athlete, "Probability of Running") +
   # ggtitle(paste("Probability of Running", stuff)) +
    theme_bw()+
    #theme(panel.border=element_rect(fill=NA)) +
    theme(axis.text=element_blank(),
          axis.title=element_blank(),
          axis.ticks=element_blank(),
          panel.grid=element_blank(),
          panel.border=element_blank())
}

val.a <- round(df_sub$gr12prob240 * 100, 0)
val.b <- round(df_sub$gr12prob235 * 100, 0)
val.c <- round(df_sub$gr12prob230 * 100, 0)
val.d <- round(df_sub$gr12prob225 * 100, 0)
val.e <- round(df_sub$gr12prob220 * 100, 0)
val.f <- round(df_sub$gr12prob215 * 100, 0)
val.g <- round(df_sub$gr12prob210 * 100, 0)
p1 <- gg.gauge_percent(val.a , breaks=c(0,30,70,100))
p2 <- gg.gauge_percent(val.b , breaks=c(0,30,70,100))
p3 <- gg.gauge_percent(val.c , breaks=c(0,30,70,100))
p4 <- gg.gauge_percent(val.d , breaks=c(0,30,70,100))
p5 <- gg.gauge_percent(val.e , breaks=c(0,30,70,100))
p6 <- gg.gauge_percent(val.f , breaks=c(0,30,70,100))
p7 <- gg.gauge_percent(val.g , breaks=c(0,30,70,100))

bang <- times * 60
bong <- round(as.hms(bang, 2))
#as.hms(bong)[7]

my.gages <- grid.arrange(arrangeGrob(p1, top = paste("Probability of", as.hms(bong)[7])),
             arrangeGrob(p2, top = paste("Probability of", as.hms(bong)[6])),
             arrangeGrob(p3, top = paste("Probability of", as.hms(bong)[5])),
             arrangeGrob(p4, top = paste("Probability of", as.hms(bong)[4])),
             arrangeGrob(p5, top = paste("Probability of", as.hms(bong)[3])),
             arrangeGrob(p6, top = paste("Probability of", as.hms(bong)[2])),
             arrangeGrob(p7, top = paste("Probability of", as.hms(bong)[1])),
             #p3, p4, p5, p6, p7, 
             #top = df_sub$Athlete,
             top = sn,
             ncol = 2)
my.gages
grid.rect(width = .98, height = .98, gp = gpar(lwd = 2, col = "black", fill = NA))
###########
```


```{r, echo = FALSE, results='hide'}
printdate <- Sys.Date()

finalprintdate <- format(printdate, format = "%B %d, %Y")

finalprintdate
```


<br>  


*****  



For information on how to subscribe and other products we offer, be sure to check us out at blahblahblah.com.

Company Name  
email address  
website  
`r finalprintdate`







```{r, echo = FALSE, warning = FALSE, message=FALSE, include = FALSE}
#like <- likelihood %>% 
#  filter(Athlete == "Lydia Alig") %>% 
#  mutate(SophPr230 = round(gr10prob230, digits = 4)) %>% 
#  select(SophPr230)

#llike <- paste(like * 100, "%", sep = "")

#llike[1]

```







